Between Subgraph Isomorphism and Maximum Common Subgraph

### 2016-09-13 paper/paper.tex
example text added.

\theoremstyle{definition}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}
\newtheorem{example}{Example}

@@

\begin{abstract}
    When a small pattern graph does not occur inside a larger target graph, we can ask how to find
    ``as much of the pattern as possible'' inside the target graph. In general, this is known as the maximum
    common subgraph problem, which is much more computationally challenging in practice than
    subgraph isomorphism. We introduce a restricted alternative, where we ask if all but $k$
    vertices from the pattern can be found in the target graph. This allows for the development of
    slightly weakened forms of certain invariants from subgraph isomorphism which are based upon degree and
    number of paths.  We show that when $k$ is small, weakening the invariants still retains much of
    their effectiveness. We are then able to solve this problem on the standard SIP problem
    instances used to benchmark subgraph isomorphism algorithms, despite these instances being
    too large for current maximum common subgraph algorithms to handle. Finally, by iteratively
    increasing $k$, we obtain an algorithm which is also competitive for the maximum common subgraph
    problem.
\end{abstract}

\section{Introduction}
\todo{talk about subgraph isomorphism applications, particularly vision}

The subgraph isomorphism problem is to find a copy of a small \emph{pattern} graph inside a larger
\emph{target} graph. It comes in two forms: in the non-induced variant, edges must be mapped to
edges, but the target may have ``extra edges'', whilst in the induced variant, non-edges may only be
mapped to non-edges. When a pattern cannot be found, we may wish to be given a result which maps as
many vertices of the pattern into the target as possible. In the induced case, this is known as the
maximum common induced subgraph problem (we discuss the non-induced case below). However, although
recent subgraph isomorphism algorithms are comfortable working with graphs with thousands of
vertices, the state of the art for the maximum common subgraph problem becomes computationally
infeasible at only 35 vertices when working with unlabelled graphs
\citep{DBLP:conf/cp/McCreeshNPS16}. This is largely because strong inference, based upon the degrees
of vertices \citep{DBLP:journals/ai/Solnon10} and the distances or paths between them
\citep{DBLP:conf/cp/AudemardLMGP14,DBLP:conf/cp/McCreeshP15}, is possible with subgraph isomorphism,
but not maximum common subgraph, and so the state space in the former is much more restricted,
whilst filtering during search is vastly stronger.

In this work we discuss an intermediate problem, where we must map all but $k$ vertices of the
pattern graph into the target. We show that if $k$ is reasonably small (say, between 1 and 5), then
weakened forms of degree and path based filterings are still effective in pruning the initial search
space and in providing additional constraints respectively. We then show that combining these
techniques leads to a practical algorithm which can scale to work with the families of graphs
commonly used to benchmark subgraph isomorphism algorithms: depending upon the benchmark family, we
can close a substantial portion of the instances, and in many more cases, we can at least obtain an
upper bound. This is a significant improvement over state of the art maximum common subgraph
solvers, which cannot even fit many of these instances in RAM. Finally, we show that iteratively
increasing $k$ gives a competitive algorithm for the maximum common induced subgraph problem.

\begin{figure}
    \begin{center}
        \begin{tikzpicture}[every node/.style={circle, inner sep=0.5mm,node distance=10mm,fill=black}]
            \node[fill=uofgcobalt] (a) {};
            \node (a1) [below left of = a] {};
            \node (b1) [below right of = a] {};
            \node (aa) [below right of = a1] {};

@@

The \emph{complement} of a (simple) graph $P$ is the graph $\shortoverline{P}~$.  Adjacent distinct
vertices in $P$ are non-adjacent in $\shortoverline{P}$ while non-adjacent vertices in $P$ are
adjacent in $\shortoverline{P}$.  The \emph{loop complement} of a graph $P$ is the graph
$\loopcomp{P}$.  The construction of $\loopcomp{P}$ is similar to that of $\bar{P}~$, however
whenever a vertex has a loop in $P$ it does not in $\loopcomp{P}$ and vice versa.  When solving an
induced subgraph isomorphism problem we make use of the loop complement; we care about mapping
vertices with loops to vertices with loops, and vertices without loops to vertices without loops.
The following proposition follows directly from the definitions.

@@

Constraints are used to ensure that adjacency pairs of vertices are mapped to adjacenct pairs of
vertices. For the maximum common subgraph case, any pair of assignments with $\bot$ as either value
is also permitted.  To handle the induced case, rather than the more usual approach of directly
having constraints for non-adjacent pairs of vertices, we will make use of \cref{prop:comp} by
seeking an $i$ which simultaneously satisfies $P \rightarrowtail T$ and $\loopcomp{P}
\rightarrowtail \loopcomp{T}$. Finally, an all different except $\bot$ constraint ensures
injectivity.

For subgraph isomorphism, this model can be enhanced with domain filtering at the top of search, to
reduce the initial sizes of domains---we discuss this in \cref{section:degreefiltering}. Subgraph
isomorphism also allows us to generate additional implied constraints, which we discuss in
\cref{section:pathfiltering}. Neither of these techniques are valid for the maximum common subgraph
problem.

@@

We perform our experiments on systems with dual Intel Xeon CPU E5-2640 processors with 64GBytes of
RAM, running Ubuntu 14.04. Our algorithm was implemented in C++\footnote{URL removed for anonymous
review}, and source code was compiled using
GCC 5.3.0. We used a timeout of 1,000 seconds for each instance. For comparison, we used the
implementations of \citet{DBLP:conf/cp/McCreeshNPS16}.

For datasets, we use the same 5,725 instances used in a recent work on portfolios of algorithms for
subgraph isomorphism \cite{thelionpaper}. This collection includes randomly generated scale-free
graphs \cite{constraints10}, an assortment of real-world graphs of varying sizes \cite{LV02},
randomly generated graph pairs (using bounded degree, regular mesh, and uniform models; all are
satisfiable), segmented images~\cite{pr15,cviu11}, meshes from modelling 3D objects \cite{cviu11},
and graphs close to the satisfiable-unsatisfiable phase transition
\cite{DBLP:conf/ijcai/McCreeshPT16}; all these instances are available in a simple text
format\footnote{\texttt{http://liris.cnrs.fr/csolnon/SIP.html}}.  Note that many of these instances
are \emph{much} larger than were used in the recent comparison of maximum common subgraph algorithms
by \citet{DBLP:conf/cp/McCreeshNPS16}: for unlabelled graphs, they used pairs of graphs with up to
35 vertices in each, whilst our dataset contains graphs with up to 6,671 vertices. For the CP
algorithm, 1,560 of the instances cannot fit in the amount of RAM we have available---we treat these
instances as having timed out. The situation for the clique encoding is even worse, and 3,653 of
these instances do not fit in RAM. In contrast, for our algorithm, every instance fits comfortably
(at least on modern server-class hardware).

@@

In \cref{figure:ids} we show that this is effective in practice: we show the amount of domain reduction
we can achieve at the top of search by using invariants and a fixed $k$, compared to the search
space size for the maximum common induced subgraph problem. (The results on the non-induced version
show a similar but slightly weaker trend.)

Some other invariants do not translate. For example, another rule which can be effective on regular
graphs involves count the number of neighbours of a vertex which are present in a triangle
\cite{mckay2014practical}. Removing a single vertex can alter this count arbitrarily, so we cannot
make use of this fact.

@@

As well as reasoning about degrees, we can also reason about paths.  A \emph{path} between two
vertices $p$ and $q$ is a sequence of edges which can be traversed from $p$ to reach $q$. We define
$\paths(p,q,n)$ to be the number of paths of length $n$ between the vertices $p$ and $q$.

@@

To allow for fast propagation, rather than calculating paths dynamically like
\citet{DBLP:conf/cp/AudemardLMGP14}, we follow the approach of \citet{DBLP:conf/cp/McCreeshP15} and
construct \emph{supplemental graphs}, and find a mapping which is simultaneously a non-induced subgraph
isomorphism between every supplemental graph pair. We use paths of length 2, looking at counts of 1,
2, and 3 in the target graph (and so counts of $1 + k$ up to $3 + k$ in the pattern).  We then
investigate whether this leads to new constraints being generated.

@@

For $k = 0$, we see many points above the $x-y$ diagonal, which shows that for many instances, we
are able to create a substantial number of new constraints at the top of search; on the other hand,
there are also points on the diagonal, which shows that sometimes this technique provides no
benefit. (Occasionally, points fall below the $x-y$ diagonal.  This is because the use of
neighbourhood degree sequence reasoning on supplemental graphs can also lead to increased domain
filtering, which could in turn eliminate a higher proportion of forbidden than permitted assignment
pairs.) For $k = 1$ and $k = 2$, the proportion of points above the diagonal diminishes, but we are
still able to create new constraints for many instances. By the time $k = 3$, most of the benefit is
disappearing---although sometimes we are still able to make a difference, and bear in mind that
sometimes adding just one new constrained pair can vastly reduce the search space.

>>>>

\theoremstyle{plain}
\newtheorem{example}{Example}

\theoremstyle{definition}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}

@@

\begin{abstract}
    When a small pattern graph does not occur inside a larger target graph, we can ask how to find
    ``as much of the pattern as possible'' inside the target graph. In general, this is known as the maximum
    common subgraph problem, which is much more computationally challenging in practice than
    subgraph isomorphism. We introduce a restricted alternative, where we ask if all but $k$
    vertices from the pattern can be found in the target graph. This allows for the development of
    slightly weakened forms of certain invariants from subgraph isomorphism which are based upon degree and
    number of paths.  We show that when $k$ is small, weakening the invariants still retains much of
    their effectiveness. We are then able to solve this problem on the standard SIP problem
    instances used to benchmark subgraph isomorphism algorithms, despite these instances being
    too large for current maximum common subgraph algorithms to handle.
\end{abstract}

\section{Introduction}
\todo{talk about subgraph isomorphism applications, particularly vision}
\todo{talk about reformulating}
The subgraph isomorphism problem is to find a copy of a small \emph{pattern} graph inside a larger
\emph{target} graph. It comes in two forms: in the non-induced variant, edges must be mapped to
edges, but the target may have ``extra edges'', whilst in the induced variant, non-edges may only be
mapped to non-edges. When a pattern cannot be found, we may wish to be given a result which maps as
many vertices of the pattern into the target as possible. In the induced case, this is known as the
maximum common induced subgraph problem (we discuss the non-induced case below). However, although
recent subgraph isomorphism algorithms are comfortable working with graphs with thousands of
vertices, the state of the art for the maximum common subgraph problem becomes computationally
infeasible at only 35 vertices when working with unlabelled graphs
\citep{DBLP:conf/cp/McCreeshNPS16}. This is largely because strong inference, based upon the degrees
of vertices \citep{DBLP:journals/ai/Solnon10} and the distances or paths between them
\citep{DBLP:conf/cp/AudemardLMGP14,DBLP:conf/cp/McCreeshP15}, is possible with subgraph isomorphism,
but not maximum common subgraph, and so the state space in the former is much more restricted,
whilst filtering during search is vastly stronger.

In this work we discuss an intermediate problem, where we must map all but $k$ vertices of the
pattern graph into the target. We show that if $k$ is reasonably small (say, between 1 and 5), then
weakened forms of degree and path based filterings are still effective in pruning the initial search
space and in providing additional constraints respectively. We then show that combining these
techniques leads to a practical algorithm which can scale to work with the families of graphs
commonly used to benchmark subgraph isomorphism algorithms: depending upon the benchmark family, in
a substantial portion of ??; in ?? more cases, we can at least obtain an upper bound. This is a
significant improvement over state of the art maximum common subgraph solvers, which cannot even fit
many of these instances in RAM.
\begin{figure}
    \begin{center}
        \begin{tikzpicture}[every node/.style={circle, inner sep=0.5mm,node distance=10mm,fill=black}]
            \node[fill=uofgcobalt] (a) {};
            \node (a1) [below left of = a] {};
            \node (b1) [below right of = a] {};
            \node (aa) [below right of = a1] {};

@@

The \emph{complement} of a (simple) graph $P$ is the graph $\shortoverline{P}~$.  Adjacent distinct
vertices in $P$ are non-adjacent in $\shortoverline{P}$ while non-adjacent vertices in $P$ are
adjacent in $\shortoverline{P}$.  The \emph{loop complement} of a graph $P$ is the graph
$\loopcomp{P}$.  The construction of $\loopcomp{P}$ is similar to that of $\bar{P}~$, however
whenever a vertex has a loop in $P$ it does not in $\loopcomp{P}$ and vice versa.  When solving an
induced subgraph isomorphism problem we make use of the loop complement; we care about mapping
vertices with loops to vertices with loops, and vertices without loops to vertices without loops.
(Although when we filter by paths we make use of the complement rather than the loop complement.)
The following proposition follows directly from the definitions.

@@

Constraints are used to ensure that adjacency pairs of vertices are mapped to adjacenct pairs of
vertices. For the maximum common subgraph case, any pair of assignments with $\bot$ as either value
is also permitted.  To handle the induced case, rather than the more usual approach of directly
having constraints for non-adjacent pairs of vertices, we will make use of \cref{prop:comp} by
seeking an $i$ which simultaneously satisfies $P \rightarrowtail T$ and $\loopcomp{P}
\rightarrowtail \loopcomp{T}$.

?? All different.

For subgraph isomorphism, we can do domain filtering at the top of search, to reduce the initial
sizes of domains---we discuss this in \cref{section:degreefiltering}. Subgraph isomorphism also
allows us to generate additional implied constraints, which we discuss in
\cref{section:pathfiltering}. Neither of these techniques are valid for the maximum common subgraph
problem.

@@

We perform our experiments on systems with dual Intel Xeon CPU E5-2640 processors with 64GBytes of
RAM, running Ubuntu 14.04. Our algorithm was implemented in C++\footnote{URL removed for anonymous
review}, and source code was compiled using
GCC 5.3.0. We used a timeout of 1,000 seconds for each instance.

For datasets, we use the same 5,725 instances used in a recent work on portfolios of algorithms for
subgraph isomorphism \cite{thelionpaper}. This collection includes randomly generated scale-free
graphs \cite{constraints10}, an assortment of real-world graphs of varying sizes \cite{LV02},
randomly generated graph pairs (using bounded degree, regular mesh, and uniform models; all are
satisfiable), segmented images~\cite{pr15,cviu11}, meshes from modelling 3D objects \cite{cviu11},
and graphs close to the satisfiable-unsatisfiable phase transition
\cite{DBLP:conf/ijcai/McCreeshPT16}; all these instances are available in a simple text
format\footnote{\texttt{http://liris.cnrs.fr/csolnon/SIP.html}}.

?? Using the French code, and the Glasgow clique code.

Note that many of these instances are \emph{much} larger than were used in the recent comparison of
maximum common subgraph algorithms by \citet{DBLP:conf/cp/McCreeshNPS16}: for unlabelled graphs,
they used pairs of graphs with up to 35 vertices in each, whilst our dataset contains patterns with
up to ?? vertices and targets with up to ?? vertices. For maximum common subgraph, 1,560 of the
instances cannot fit in the amount of RAM we have available---we treat these instances as having
timed out. The situation for the clique encoding is even worse, and 3,653 of these instances do not
fit in RAM. In contrast, for our algorithm, every instance fits comfortably (at least on modern
server-class hardware).

@@

In \cref{figure:ids} we show that this is useful in practice.

Some other forms do not translate. For example, another invariant which can be effective on regular
graphs is to count the number of neighbours of a vertex which are present in a triangle \cite{mckay2014practical}. Removing a single vertex can alter this count arbitrarily, so we cannot make use of this fact.

@@

As well as reasoning about degrees, we can also reason about paths.

A \emph{path} between two vertices $p$ and $q$ is a sequence of edges which can be traversed from $p$ to reach $q$. We define $\paths(p,q,n)$ to be the number of paths of length $n$ between the vertices $p$ and $q$.
%\todo{Define paths}

@@

To allow for fast propagation, rather than calculating paths dynamically, we follow the approach of
\citet{DBLP:conf/cp/McCreeshP15} and construct supplemental graphs, and find a mapping which is
simultaneously a non-induced subgraph isomorphism between every supplemental graph pair. We use
paths of length 2, looking at counts of 1, 2, and 3 in the target graph (and so counts of $1 + k$ up
to $3 + k$ in the pattern).  We then investigate whether this leads to new constraints being
generated.

@@

For $k = 0$, we see many points above the $x-y$ diagonal, which shows that for many instances, we
are able to create a substantial number of new constraints at the top of search; on the other hand,
there are also points on the diagonal, which shows that sometimes this technique provides no
benefit. (Occasionally, points fall below the $x-y$ diagonal.  This is because the use of
neighbourhood degree sequence reasoning on supplemental graphs can also lead to increased domain
filtering, which could in turn eliminate a higher proportion of forbidden than permitted assignment
pairs.) For $k = 1$ and $k = 2$, the proportion of points above the diagonal diminishes, but we are
still able to create new constraints for many instances. By the time $k = 3$, most of the benefit is
disappearing---although sometimes we are still able to make a difference, and bear in mind that
sometimes adding just one new constrained pair can vastly reduce the search space.  A closer
inspection of the raw data shows that for $k \ge 3$, nearly all of the instances showing a strong
improvement are from the ``LV'' family of graphs.
